\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{geometry}
\geometry{margin=1in}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{booktabs}

\title{SAM3 Paper 13: Overall Rigor, Notation Consistency, Cross-References, and Unified Presentation of the Framework}
\author{Shawn Dykes \\ SAM3 Framework}
\date{v4.23 \quad May 2026}

\begin{document}

\maketitle

\begin{abstract}
This paper addresses the global mathematical rigor of the SAM3 framework. We establish consistent notation across all 13 papers, provide a master cross-reference map, update all instances of the Newton constant to the latest value \(G_N = \frac{64\pi \ell_0^2}{45}\), and present a unified summary of the extraordinary claims together with the rigorous foundations now established in Papers 09--12.
\end{abstract}

\section{Introduction}
The SAM3 series consists of 13 self-contained papers plus a merged full document. While each paper focuses on a specific component, the complete framework must be internally consistent and rigorously cross-linked. This document provides that global consistency layer.

\section{Unified Notation and Conventions}

All papers now adopt the following global conventions (updated where necessary):

\begin{itemize}
    \item \(\ell_0\): fundamental length scale of the right conoid (sets all masses and \(G_N\)).
    \item \(\omega_0 > 0\): Dual-Zero strength parameter.
    \item \(D_\infty\): Dirac operator on the infinite right conoid.
    \item \(D = D_\infty \otimes 1 + 1 \otimes D_F\): full almost-commutative Dirac operator.
    \item \(\epsilon(n) = \omega_0 (-1)^n n^{-n}\): Dual-Zero infinitesimal (Paper 09).
    \item \(G_N = \frac{64\pi \ell_0^2}{45}\): Newton constant (Papers 10 and 05, updated everywhere).
    \item \(J(k_1,k_2)\): information current (Paper 11).
    \item \(S_I = \int |J|^2 \, d\mu\): information action enforcing \(\operatorname{Re}(s)=1/2\) (Paper 11).
\end{itemize}

A master symbol list is maintained in the repository file `math/notation_master.tex` and is included via `\input` in every paper.

\section{Cross-Reference Map}

\begin{table}[ht]
\centering
\begin{tabular}{ll}
\toprule
Topic & Primary Paper(s) \\
\midrule
Right Conoid Geometry \& Metric & Paper 01, Paper 10 \\
Dual-Zero Algebra \& Information Conservation & Paper 09 \\
Dimensional Lift \& Seeley--DeWitt & Paper 10 \\
Variational RH Mechanism & Paper 11 \\
Numerical Pipeline, Convergence \& Error Bars & Paper 12 \\
Yukawa Matrices \& CKM/PMNS from Overlaps & Paper 04, Paper 12 \\
Gravity Sector \& \(G_N\) & Paper 05, Paper 10 \\
Full Spectral Triple \& Standard Model & Paper 02, Paper 03 \\
Black-Hole Extension \& Information Flow & Paper 06 \\
Unified Summary \& Extraordinary Claims & \textbf{Paper 13 (this work)} \\
\bottomrule
\end{tabular}
\caption{Master cross-reference table.}
\end{table}

Every paper now contains at the end of its introduction:  
``For global notation and cross-references see Paper 13.''

\section{Update of the Newton Constant}

All previous references to the earlier value \(G_N = \frac{3\pi \ell_0^2}{2}\) have been replaced by the corrected derivation:
\[
G_N = \frac{64\pi \ell_0^2}{45}.
\]
This update has been propagated through Papers 01, 05, 10, and the merged full paper (v4.23). The factor \(64/45\) arises from the precise angular integration over the 12-bridge icosahedral measure and the full \(f(u,v)\) metric factor (see Paper 10 for the detailed calculation).

\section{Unified Statement of the Extraordinary Claims}

The SAM3 framework asserts that a **single geometric object** — the infinite right conoid with 12 discrete bridges carrying \(2I\) symmetry — together with the Dual-Zero regularization, produces:

1. Einstein gravity in 4D with \(G_N = \frac{64\pi \ell_0^2}{45}\) (Papers 05, 10).
2. The full chiral Standard Model gauge group, fermion representations, three generations, hierarchical Yukawa matrices, and realistic CKM/PMNS mixing via geometric overlaps (Papers 02--04, 12).
3. Type-I seesaw neutrino masses and a stable Higgs vacuum (Paper 07).
4. A variational information principle that enforces the Riemann Hypothesis critical line exactly (Paper 11).
5. Information conservation at every scale via the Dual-Zero algebra (Paper 09).

These claims are no longer heuristic. Each component is now supported by:
- Rigorous proofs (dimensional lift, RH stationarity, uniqueness of Dual-Zero).
- Complete numerical documentation with error bars (Paper 12).
- Full open-source reproducibility.

\section{Independent Research Note}
SAM3 is an independent research program. External citations to noncommutative geometry, spectral action, and related literature are provided where standard techniques are used. All novel constructions (right conoid + 12 bridges, Dual-Zero \(\epsilon(n)\), information current \(J\)) are original to this framework.

\section{Conclusion and Outlook}

With the addition of Papers 09--13, the SAM3 series now meets the standards of rigor required for its unification claims. The framework is internally consistent, fully cross-referenced, numerically validated, and philosophically grounded in exact information conservation.

Future directions include phenomenological predictions beyond the Standard Model, higher-order Dual-Zero effects, and potential experimental signatures.

All code, LaTeX sources, and notebooks are available at:  
\href{https://github.com/mohawksd9sd-maker/SAM3-DualZero-Conoid}{https://github.com/mohawksd9sd-maker/SAM3-DualZero-Conoid}

\bibliographystyle{plain}
\begin{thebibliography}{13}
\bibitem{SAM3-09} Dual-Zero Algebra (Paper 09).
\bibitem{SAM3-10} Dimensional Lift (Paper 10).
\bibitem{SAM3-11} RH Variational Principle (Paper 11).
\bibitem{SAM3-12} Numerical Aspects (Paper 12).
\end{thebibliography}

\end{document}
