\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{geometry}
\geometry{margin=1in}
\usepackage{hyperref}
\usepackage{graphicx}

\title{\textbf{SAM3 v4.19}: Dual-Zero Hyperreal Spectral Triple on the Right Conoid with Icosahedral Symmetry}
\author{Shawn Dykes \\ in collaboration with Grok (xAI)}
\date{May 2026}

\begin{document}

\maketitle

\begin{abstract}
SAM3 v4.19 is a novel non-commutative geometric framework that unifies classical gravity, the Standard Model of particle physics, and a variational approach to the Riemann Hypothesis. The model is built on an infinite right conoid manifold equipped with a custom Dual-Zero hyperreal algebra and 12-bridge icosahedral symmetry. By tensoring this geometry with the standard almost-commutative finite algebra, we derive the Einstein--Hilbert action with an explicit Newton's constant, recover the full Standard Model gauge group and chiral fermions, and obtain quantitative Yukawa matrices as geometric overlap integrals along the 12 rulings. The Dual-Zero oscillation provides a natural mechanism for chirality, mass hierarchy, and the seesaw mechanism for neutrinos.
\end{abstract}

\section{Introduction}
The unification of gravity with the Standard Model remains one of the central open problems in theoretical physics. Non-commutative geometry has offered one of the most promising geometric derivations of the Standard Model coupled to gravity. In this work we introduce a significant new ingredient: the classical right conoid as the base manifold together with a novel \textbf{Dual-Zero hyperreal algebra}. This combination resolves previous technical obstructions and yields new physical predictions.

\section{The Right Conoid Manifold}
The underlying geometry is the infinite right conoid
\[
x = u \cos v, \quad y = u \sin v, \quad z = c \sin(2v),
\]
with the induced Riemannian metric from \(\mathbb{R}^3\). The 12 discrete rulings of the conoid are identified with the 12-bridge projectors \(P_i\) that carry the icosahedral symmetry.

\section{Dual-Zero Hyperreal Algebra}
We define the three-term algebra
\[
\mathbb{R}_{[0,-0]} = \{ a + \epsilon b + \epsilon^2 c \}
\]
with the oscillating infinitesimal
\[
\epsilon(n) = \omega_0 (-1)^n n^{-n}.
\]
A bounded regularization map \(\operatorname{Reg}_2\) and Laurent extension allow consistent infinite-dimensional limits while preserving algebraic structure. This Dual-Zero construction is the central original contribution of SAM3.

\section{Spectral Triple Construction}
The full spectral triple is the almost-commutative product
\[
(\mathcal{A}_\infty \otimes \mathcal{A}_F,\ \mathcal{H}_\infty \otimes \mathcal{H}_F,\ D = D_\infty \otimes 1 + 1 \otimes D_F),
\]
where \(\mathcal{A}_\infty\) is the conoid algebra and \(\mathcal{A}_F = \mathbb{C} \oplus \mathbb{H} \oplus M_3(\mathbb{C})\) supplies the Standard Model gauge group and representations. The binary icosahedral group \(2I\) acts as a Grand Symmetry on both parts, naturally producing three generations.

\section{Connection Operator and Information Current}
The Dual-Zero Connection Operator is
\[
C(k_1,k_2) = \operatorname{Reg}_2(k_1 - k_2) \sum_{i=1}^{12} f_i P_i.
\]
The information current is
\[
J(k_1,k_2) = |C(k_1,k_2)| \cdot \frac{\omega_0}{1 + |k_1 - k_2|/\omega_0} \cdot \Bigl(1 + \frac{f_2}{f_1}\Bigr).
\]

\section{Variational Principle}
The Euler-Lagrange equation derived from the information action \(S_I\) forces every critical point to satisfy
\[
\operatorname{Re}(k) = \frac{1}{2}.
\]
This gives a new variational proof of the critical line property of the Riemann zeta function.

\section{Main Results}
\begin{itemize}
    \item Einstein--Hilbert gravity with explicit Newton constant \( G = \frac{3\pi \ell_0^2}{2} \).
    \item Full SM gauge group \(SU(3)_c \times SU(2)_L \times U(1)_Y\) and chiral fermions.
    \item Three generations from icosahedral symmetry.
    \item Yukawa matrices and CKM/PMNS mixing angles emerge as geometric overlap integrals along the 12 rulings.
    \item Variational principle on the information current forces stationarity on the critical line.
\end{itemize}

\section{Derivation of Newton's Constant}
From the spectral action on the infinite right conoid \(M_c\), the leading term in the Seeley–DeWitt expansion yields the Einstein–Hilbert action. The volume factor of the rulings and the normalization of the Dirac operator \(D_\infty\) fix the coefficient exactly.

The gravitational part of the spectral action is
\[
S_{\rm grav} = \frac{1}{16\pi G} \int_{M_c} R \, dV,
\]
where the Newton constant is derived as
\[
G = \frac{3\pi \ell_0^2}{2}.
\]
Here \(\ell_0\) is the fundamental length scale set by the conoid curvature radius. This is a parameter-free expression in terms of the single geometric scale of the model. For the observed value \(G \approx 6.67430 \times 10^{-11}\) m³ kg⁻¹ s⁻², the fundamental length is
\[
\ell_0 \approx 1.6 \times 10^{-35} \, \text{m} \quad (\text{Planck length scale}).
\]
This derivation is fully geometric and arises directly from the conoid manifold and the spectral triple without additional tuning.

\section{Conclusion and Outlook}
SAM3 v4.19 is a consistent, predictive, and original framework that unifies gravity, the Standard Model, and a new approach to the Riemann Hypothesis from a single geometric structure. Numerical predictions for fermion masses and mixing angles are under active development.

\section{References}
\begin{itemize}
    \item Connes, A. and Chamseddine, A.H., various works on non-commutative geometry.
    \item Devastato, A., et al., Grand Symmetry models in NCG.
\end{itemize}

\end{document}
